On centrality of orthogonal $\Kt$

### 2016-04-25 paper.tex
More fixes

In order to construct the arrow $T$ Tulenbaev uses the following presentation for the relative Steinberg group, see~\cite[Proposition~1.6]{Tul}. 
%Assume that $I\trianglelefteq R$ such that a short exact sequence
%\renewcommand{\theequation}{$\dagger$}
%\begin{align} \xymatrix{0\ar@<0.1ex>[r]^{}&I\ar@<0.1ex>[r]^{}&R\ar@<0.5ex>[r]^{\pi}&R/I\ar@<0.5ex>[l]^{\sigma}\ar@<0.1ex>[r]^{}&0} \end{align} splits, i.e., $\pi\sigma=1$.
\begin{prop}\label{prop:TulPres}
For a splitting ideal $I$ the group $\St(n,\,R,\,I)$ is isomorphic to the group given by generators
$$\{X(u,\,v)\mid u\in\E(n,\,R)e_1,\ v\in I^n,\ u^tv=0\}$$ subject to relations
\setcounter{equation}{0}
\renewcommand{\theequation}{T\arabic{equation}}
\begin{align}
&X(u,\,v)X(u,\,w)=X(u,\,v+w), \label{add2}\\
&X(u,\,v)X(u',\,v')X(u,\,v)\inv=X(t(u,\,v)u',\,t(v,\,u)\inv v'), \label{conj2}  \\
&X(ur+w,\,v)=X(u,\,vr)X(w,\,v)\,\text{ for }r\in R,\ (u,\,w)\in\Um_{n\times2}(R) \label{add3}.
\end{align}
\end{prop}
Here $\Um_{n\times2}(R)$ denotes the set of $n\times2$ \emph{unimodular matrices}, i.\,e. matrices $M$ such that there exists a $2\times n$ matrix $N$ satisfying $NM=\begin{pmatrix}1&0\\0&1\end{pmatrix}$.

Both ``another presentations'' of van der Kallen and Tulenbaev are given in terms of generators parametrised by pairs of vectors,
 where the first one is in some sense ``good'' (i.\,e. unimodular or a column of an elementary matrix) and the second one is arbitrary.
It is easy to formulate the additivity property in the second argument for these generators (cf. \eqref{add1}, \eqref{add2}),
while it is not so easy when it comes for the additivity in the first argument (cf. \eqref{add3}).
%Such presentations allowed them to formulate easily an additivity by the second argument, but one has to write something similar to T3 instead of additivity by the first one.

The construction of the homomorphism $T\colon\St(n,\,R_a[X],\,XR_a[X])\rightarrow\St(n,\,R\ltimes XR_a[X]])$
amounts to choosing certain elements in the target group $\St(n,\,R\ltimes XR_a[X]])$ and proving that these elements satisfy the relations \eqref{add2}--\eqref{add3}, see~\cite[Lemmas~1.2 and~1.3\,c)]{Tul}.
The main problem with this recipe is that the assumption $n\geq5$ is essential to check that the relation~\eqref{add3} holds.

@@

Clearly, we must map the elements $F(u,\,v)$ to van der Kallen's generators $X(u,\,v)$.
Now we should find the images for the elements $S(u,\,v)$ as well.
First of all, recall that van der Kallen in~\cite[3.8--3.10]{vdK} constructs the following elements
$$x(u,\,v)\in\St(n,\,R),\ u^tv=0,\ u_i=0\,\text{ or }\,v_i=0\,\text{ for some }\,1\leq i\leq n.$$
The following definition is the ``transpose'' of \cite[3.13]{vdK}.
\begin{df} For $u\in R^n$, $v\in\E(n,\,R)e_1$, $u^tv=0$, consider the set $\overline Y(u,\,v)\subseteq\St(n,\,R)$
 consisiting of all elements $y\in\St(n,\,R)$ that admit the decomposition $y=\prod x(w^k,\,v)$, 
 where $\sum u^k=u$ and $u^k=c^k \cdot (e_pv_q-e_qv_p)$ for some $c^k\in R$, $1\leq p\neq q\leq n$. \end{df}

@@

We denote the only element of $\overline Y(u,\,v)$ by $Y(u,\,v)$. 
Now we can define $\iota$ by sending $S(u,\,v)\in\St^*(n,\,R,\,I)$ to $Y(u,\,v)$. 
To show that the map $\iota$ is well-defined we should check that the elements $X(u,\,v)$ and $Y(u,\,v)$ satisfy the relations 
\eqref{add4}--\eqref{coinc} (with $F$'s and $S$'s replaced with $X$'s and $Y$'s).
Notice that the validity of relations \eqref{add4}--\eqref{conj3} has already been checked.
\begin{lm} For $(u,\,v^t)=(Me_1,\,e_2^tM\inv)$, $M\in\E(n,\,R)$, $a\in I$ one has $X(u,\,va)=Y(ua,\,v)$. \end{lm}
\begin{proof}
In view of the above lemma we only need to show that $X(e_1,\,e_2a)=Y(e_1a,\,e_2)$.
This equality can be obtained by computing the commutator $[Y(-e_3,\,e_2),\,X(e_1,\,e_3a)]$ in two ways:
$$ Y(-e_3,\,e_2)X(e_1,\,e_3a)Y(-e_3,\,e_2)\inv\cdot X(e_1,\,-e_3a)=X(e_1,\,e_2a), \text{ and} $$
$$ Y(-e_3,\,e_2)\cdot\,X(e_1,\,e_3a)Y(e_3,\,e_2)X(e_1,\,e_3a)\inv=Y(e_1a,\,e_2). \qedhere $$
\end{proof}

We can sum up the above discussion into the following result.
\begin{lm} \label{lm:map-iota}
 For any $I \trianglelefteq R$ there exists a well-defined homomorphism $$\iota\colon\St^*(n,\,R,\,I)\rightarrow\St(n,\,R, I) \subseteq \St(n, R)$$ 
 satisfying the equalities $\iota (F(u,\,v)) = X(u,\,v)$ and $\iota (S(u,\,v)) = Y(u,\,v)$. \end{lm}
\begin{proof}
We already know that $\iota$ is well-defined. It suffices to check that the image of $\iota$ is contained in $\St(n,\,R,I)$.
Denote the canonical projection $R\epi R/I$ by $\rho$.
Clearly, the induced group morphism $\rho^*\colon\St(n,\,R)\rightarrow\St(n,\,R/I)$ sends $X(u,\,v)$ to $X(\rho(u),\,\rho(v))$ and $Y(u,\,v)$ to $Y(\rho(u),\,\rho(v))$. 
Observe that $\E(n,\,R)\epi\E(n,\,R/I)$ is surjective, therefore for $x\in\E(n,\,R)e_1$ one has $\rho(x)\in\E(n,\,R/I)e_1$. 
From the additivity relations we get that $X(u,\,0)=1=Y(0,\,v)$, hence $\mathrm{Im}(\iota)\subseteq\Ker(\rho^*)=\St(n,\,R,\,I)$, as required. \end{proof}

We constructed $\iota$ for an arbitrary $I\trianglelefteq R$. For $I$ a splitting ideal we construct an inverse homomorphism
$$
\kappa\colon\St(n,\,R,\,I)\rightarrow\St^*(n,\,R,\,I).
$$
With this end we recall a presentation of $\St(n,\,R,\,I)$ as a group with action of $\St(n,\,R)$. For a splitting ideal such a presentation was firstly obtained by Swan~\cite{Swa1,Swa2,Swa3}, and for an arbitrary $I$ by Keune~\cite{Keu} and Loday~\cite{Lod}.

@@

Next, define $y_{ij}^*(a)=F(e_i,\,e_ja)$, $a\in I$. These elements obviously satisfy KL1, KL2 and KL5. Check KL3:
\begin{multline*}
\!\,^{x_{ij}(r)}y^*_{jk}(b)=F(t_{ij}(r)e_j,\,t_{ji}(-r)e_kb)=F(e_ir+e_j,\,e_kb)=\\
=S(e_irb+e_jb,\,e_k)=S(e_irb,\,e_k)S(e_jb,\,e_k)=F(e_i,\,e_krb)F(e_j,\,e_kb).
\end{multline*} 
KL4 is similar. Finally, we have a well-defined map
$$
\kappa\colon\St(n,\,R,\,I)\rightarrow\St^*(n,\,R,\,I)
$$
sending $y_{ij}(a)$ to $y^*_{ij}(a)$. Obviously, $\iota\circ\kappa=\mathrm{id}$ (cf.~\cite[3.6\,d)]{vdK}), thus $\kappa$ is injective. Surjectivity of $\kappa$ follows from Lemma~\ref{allyouneedisf}\,a), thus it is an isomorphism and $\iota$ is inverse to it.

>>>>

In order to construct the arrow $T$ Tulenbaev uses the following presentation for the relative Steinberg group, see~\cite[Proposition~1.6]{Tul}. 
%Assume that $I\trianglelefteq R$ such that a short exact sequence
%\renewcommand{\theequation}{$\dagger$}
%\begin{align} \xymatrix{0\ar@<0.1ex>[r]^{}&I\ar@<0.1ex>[r]^{}&R\ar@<0.5ex>[r]^{\pi}&R/I\ar@<0.5ex>[l]^{\sigma}\ar@<0.1ex>[r]^{}&0} \end{align} splits, i.e., $\pi\sigma=1$.
\begin{prop}\label{prop:TulPres}
For a splitting ideal $I$ and $n\geq 4$ the group $\St(n,\,R,\,I)$ is isomorphic to the group defined by generators
$$\{X(u,\,v)\mid u\in\E(n,\,R)e_1,\ v\in I^n,\ u^tv=0\}$$ and the following set of relations
\setcounter{equation}{0}
\renewcommand{\theequation}{T\arabic{equation}}
\begin{align}
&X(u,\,v)X(u,\,w)=X(u,\,v+w), \label{add2}\\
&X(u,\,v)X(u',\,v')X(u,\,v)\inv=X(t(u,\,v)u',\,t(v,\,u)\inv v'), \label{conj2}  \\
&X(ur+w,\,v)=X(u,\,vr)X(w,\,v)\,\text{ for }r\in R,\ (u,\,w)\in\Um_{n\times2}(R) \label{add3}.
\end{align}
Moreover, the group $\St(n, R)$ decomposes into the semidirect product $\St(n,\,R, I)\rtimes \St(n,\,R/I)$.
\end{prop}
Here $\Um_{n\times2}(R)$ denotes the set of $n\times2$ \emph{unimodular matrices}, i.\,e. matrices $M$ such that there exists a $2\times n$ matrix $N$ satisfying $NM=\begin{pmatrix}1&0\\0&1\end{pmatrix}$.

Both ``another presentations'' of van der Kallen and Tulenbaev are given in terms of generators parametrised by pairs of vectors,
 where the first one is ``good'' in some sense  (i.\,e. unimodular or a column of an elementary matrix) while the second one is arbitrary.
It is easy to formulate the additivity property in the second argument for these generators (cf. \eqref{add1}, \eqref{add2}),
while it is not so easy when it comes to the additivity in the first argument (cf. \eqref{add3}).
%Such presentations allowed them to formulate easily an additivity by the second argument, but one has to write something similar to T3 instead of additivity by the first one.

The construction of the homomorphism $T\colon\St(n,\,R_a[X],\,XR_a[X])\rightarrow\St(n,\,R\ltimes XR_a[X])$
amounts to choosing certain elements in the group $\St(n,\,R\ltimes XR_a[X])$ and proving that these elements satisfy relations \eqref{add2}--\eqref{add3}, see~\cite[Lemmas~1.2 and~1.3\,c)]{Tul}.
The main problem with this recipe is that the assumption $n\geq5$ is essential to check that the relation~\eqref{add3} holds.

@@

Clearly, we must map the elements $F(u,\,v)$ to van der Kallen's generators $X(u,\,v)$.
Now we should find the images for elements $S(u,\,v)$ as well.
First of all, recall that van der Kallen in~\cite[3.8--3.10]{vdK} constructs the following elements
$$x(u,\,v)\in\St(n,\,R),\ u^tv=0,\ u_i=0\,\text{ or }\,v_i=0\,\text{ for some }\,1\leq i\leq n.$$
The following definition is the ``transpose'' of \cite[3.13]{vdK}.
\begin{df} For $u\in R^n$, $v\in\E(n,\,R)e_1$, $u^tv=0$, consider the set $\overline Y(u,\,v)\subseteq\St(n,\,R)$
 consisiting of all elements $y\in\St(n,\,R)$ that admit the decomposition $y=\prod x(w^k,\,v)$, 
 where $\sum u^k=u$ and $u^k=c^k \cdot (e_pv_q-e_qv_p)$ for some $c^k\in R$, $1\leq p\neq q\leq n$. \end{df}

@@

We denote the only element of $\overline Y(u,\,v)$ by $Y(u,\,v)$. 
Now we can finish the definition of $\iota$ by requiring that $\iota(S(u,\,v)) = Y(u,\,v)$ for $S(u,\,v)\in\St^*(n,\,R,\,I)$. 
To show that the map $\iota$ is well-defined we should check that elements $X(u,\,v)$ and $Y(u,\,v)$ satisfy the relations 
\eqref{add4}--\eqref{coinc} (with $F$'s and $S$'s replaced with $X$'s and $Y$'s).
Only the validity of relation \eqref{coinc} is not immediately obvious.
\begin{lm} \label{lm:XY} For $(u,\,v^t)=(Me_1,\,e_2^tM\inv)$, $M\in\E(n,\,R)$, $a\in I$ one has $X(u,\,va)=Y(ua,\,v)$. \end{lm}
\begin{proof}
In view of the above lemma we only need to show that $X(e_1,\,e_2a)=Y(e_1a,\,e_2)$.
This equality can be obtained by computing the commutator $[Y(-e_3,\,e_2),\,X(e_1,\,e_3a)]$ in two ways:
$$ Y(-e_3,\,e_2)X(e_1,\,e_3a)Y(-e_3,\,e_2)\inv\cdot X(e_1,\,-e_3a)=X(e_1,\,e_2a), \text{ and} $$
$$ Y(-e_3,\,e_2)\cdot\,X(e_1,\,e_3a)Y(e_3,\,e_2)X(e_1,\,e_3a)\inv=Y(e_1a,\,e_2). \qedhere $$
\end{proof}

The following result is a consequence of the above argument.
\begin{lm} \label{lm:map-iota}
 For any $I \trianglelefteq R$ there exists a well-defined homomorphism $$\iota\colon\St^*(n,\,R,\,I)\rightarrow\St(n,\,R, I) \subseteq \St(n, R)$$ 
 satisfying the equalities $\iota (F(u,\,v)) = X(u,\,v)$ and $\iota (S(u,\,v)) = Y(u,\,v)$. \end{lm}
\begin{proof}
We already know that $\iota$ is well-defined. It suffices to check that the image of $\iota$ is contained in $\St(n,\,R,I)$.
Denote the canonical projection $R\epi R/I$ by $\rho$.
Clearly, the induced group morphism $\rho^*\colon\St(n,\,R)\rightarrow\St(n,\,R/I)$ sends $X(u,\,v)$ to $X(\rho(u),\,\rho(v))$ and $Y(u,\,v)$ to $Y(\rho(u),\,\rho(v))$. 
Observe that $\E(n,\,R)\epi\E(n,\,R/I)$ is surjective, therefore for $x\in\E(n,\,R)e_1$ one has $\rho(x)\in\E(n,\,R/I)e_1$. 
From the additivity relations we get that $X(u,\,0)=1=Y(0,\,v)$, hence $\mathrm{Im}(\iota)\subseteq\Ker(\rho^*)=\St(n,\,R,\,I)$, as required. \end{proof}

In order to construct the arrow $\kappa$ we use the presentation from \cref{prop:TulPres}.
We set $$\kappa(X(u,v)) = F(u,v),\ u\in \E(n, R)e_1,\ v\in I^n.$$ 
We have to check that $F(u,v)$ satisfy the relations \eqref{add2}--\eqref{add3}.
The relations \eqref{add2} and \eqref{conj2} are immediately obvious.

\begin{comment}
We constructed $\iota$ for an arbitrary $I\trianglelefteq R$. For $I$ a splitting ideal we construct an inverse homomorphism
$$
\kappa\colon\St(n,\,R,\,I)\rightarrow\St^*(n,\,R,\,I).
$$
With this end we recall a presentation of $\St(n,\,R,\,I)$ as a group with action of $\St(n,\,R)$. For a splitting ideal such a presentation was firstly obtained by Swan~\cite{Swa1,Swa2,Swa3}, and for an arbitrary $I$ by Keune~\cite{Keu} and Loday~\cite{Lod}.

@@

Next, define $y_{ij}^*(a)=F(e_i,\,e_ja)$, $a\in I$. These elements obviously satisfy KL1, KL2 and KL5. Check KL3:
\begin{multline*}
\!\,^{x_{ij}(r)}y^*_{jk}(b)=F(t_{ij}(r)e_j,\,t_{ji}(-r)e_kb)=F(e_ir+e_j,\,e_kb)=\\
=S(e_irb+e_jb,\,e_k)=S(e_irb,\,e_k)S(e_jb,\,e_k)=F(e_i,\,e_krb)F(e_j,\,e_kb).
\end{multline*} 
KL4 is similar. Finally, we have a well-defined map
$$
\kappa\colon\St(n,\,R,\,I)\rightarrow\St^*(n,\,R,\,I)
$$
sending $y_{ij}(a)$ to $y^*_{ij}(a)$. Obviously, $\iota\circ\kappa=\mathrm{id}$ (cf.~\cite[3.6\,d)]{vdK}), thus $\kappa$ is injective. Surjectivity of $\kappa$ follows from Lemma~\ref{allyouneedisf}\,a), thus it is an isomorphism and $\iota$ is inverse to it.
\end{comment}
